---
title: "HW2: Markdown file"
author: "Kevin Liu"
date: "2023-09-30"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

## First, clean the data in pols-month.csv. 

Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r}
pols_month_df = 
  read_csv("./data/pols-month.csv") |>
  janitor::clean_names() |> 
  separate(mon, into = c("year", "month", "day"), convert = TRUE) |> 
  mutate(month = month.abb[month]) |> 
  
  #This leaves leading zeros, which int will delete... but should be fine...
  mutate(year = substr(year,2,3)) |>  
  mutate(year = as.integer(year)) |> 
#Redundant code...I should have used "convert"!
  # separate(mon, into = c("year", "month_day"), sep = 4) |> 
  # mutate(year = as.numeric(year)) |> 
  # separate(month_day, into = c("month", "day"), sep = 3) |> 
  # mutate(month = substr(month, 2,3)) |> 
  # mutate(month = as.numeric(month)) |> 
  # mutate(month = month.name[month]) |> 
  # mutate(day = substr(day, 2,3)) |> 
  # mutate(day = as.numeric(day)) |> 

  mutate(president = if_else(prez_gop == 1 | prez_gop == 2, "gop", "dem")) |> 
  select(year, everything()) |> 
  select(-prez_gop, -prez_dem, -day)
#pols_month_df
```


## Second, clean the data in snp.csv using a similar process to the above. 

For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp_df = 
  read_csv("./data/snp.csv") |> 
  janitor::clean_names() |> 
  separate(date, into = c("month", "day", "year"), convert = TRUE) |> 
  mutate(month = month.abb[month]) |>
  select(-day) |> 
  select(year,everything())

#The way i did is not robust for the given date
  # separate(date, into = c("month", "day_year"), sep = 1) |> 
  # mutate(month = as.numeric(month)) |> 
  # mutate(month = month.name[month]) |> 
  # separate(day_year, into = c("day", "year"), sep = 2) |> 
  # select(-day) |> 
  # mutate(year = substr(year, 2,3)) |> 
  # mutate(year = if_else(grepl("/", year), substr(year,2,3), year))
  # mutate(year = as.numeric(year)) |> 
  # arrange(year)
#snp_df
```

## Third, tidy the unemployment data so that it can be merged with the previous datasets. 

This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}
unemployment_df = 
  read_csv("./data/unemployment.csv") |> 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment") |> 
  mutate(Year = substr(Year,3,4)) |> 
  rename("year" = "Year") |> 
  mutate(year = as.integer(year))
```

## Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
polls_snp_df = 
  left_join(pols_month_df, snp_df)

polls_snp_unem_df =
  left_join(polls_snp_df, unemployment_df)
```

### Write a short paragraph about these datasets. 

Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

__Response__
  The "pols" dataset contains 822 observations of 9 variables related to the number of national politicians who are democratic or republican at any given time from 1947 to 2015. Of the 9 variables, 1 is the date, 4 variables enumerate how many presidents, govenors, senators, and representatives that are democrats, and the last 4 is the same enumeration but for repbulicans.
  The "snp" dataset contains 787 observations of 2 variables related to Standard & Poor’s stock market index (S&P). It contains a date variable and the S&P's closing value on that date from 1950 to 2015.
  The "unemployment" dataset contains 68 observations of 13 variables. The first column variable the year, denotes the year for the row while the next 12 column variables designates the month of that year. Within the cells are the unemployment rates for a partcular month in a year. During data processing, "unemployement" was long-ified and contains data from 1948 to 2015.
  The resulting dataset has a combination of all three datasets. "Unemployment" has been made long-ified in order to be properly merged with "pols" and "snp" based on month and year. It is a dataset with 822 observations and 11 variables where all three were merged based on year + month with close + unemployment being added to "pols" dataset.
  
# Problem 2

Read and clean the Mr. Trash Wheel sheet:

*   specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures;  columns containing notes) using arguments in read_excel
*   use reasonable variable names
*   omit rows that do not include dumpster-specific data

```{r}
trash_df =
  readxl::read_excel()
```

